Fixing Vulnerabilities Hinders Maintainability

### 2019-02-03 paper/ICPC19/icpc19.tex
charts fixed
more fixes

In this paper, our concern is to study whether, while improving software
security, developers are also negatively impacting the maintainability
of their software application. To answer the following two
research questions, we use a dataset of security refactorings to measure the
impact of the changes in the maintainability of open-source software.

@@

Software with \emph{Broken Authentication} flaws are the most affected ones by security
refactorings. 55.3\% (21) of the refactorings had a negative impact in the maintainability
while 35\% (11) yielded a positive impact and only 13.8\% (6) did not suffer any
change. The results also show that fixing \emph{Memory Leak}s affects the maintainability
of 51.3\% (41) of the cases and leaves only 13.8\% (11) of the cases unaffected.
In 35\% (28) of the cases, maintainability increased. Refactoring vulnerabilities
that allow \emph{Denial-of-Dervice} attacks (e.g., CVE-2014-3506) has a negative impact
of 47.5\% (19) in the maintainability of open-source software. 30.0\% (12) of
the cases exhibit maintainability improvement while 22.5\% (9) show that overall
maintainability does not suffer any change. \emph{Miscellaneous} refactorings pattern
integrates several types of security flaws whose impact is of 43.1\% (84) negative
in the maintainability of open-source sofware. 31.8\% (62) of the cases do not
exhibit any change in maintainability and 25.1\% (49) of cases have a positive
impact on the maintainability of open-source software.

It is also important to notice that overall the maintainability of open-source
software is not affected by \emph{Cross-Site Scripting} vulnerabilities and by \emph{Using
Components with Known Vulnerabilities}. \emph{Cross-Site Scripting} does not suffer any
decrease in the maintainability of 55.3\% (63) of cases and 21.1\% (24) have a
positive effect on it. Only 23.7\% (27) of the cases hinder the software
maintainability. From the 21 cases evaluated for \emph{Using Componentes with Known
Vulnerabilities}, only 4.8\% (1) has a positive effect in the maintainability
of software. 14.3\% (3) harms maintainability software and 81\% has no effect
in the maintainability.

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.55\textwidth]{figures/category.pdf}
 	\caption{Maintainability difference by pattern after security refactorings}
	\label{fig:pat}
\end{figure}

@@

We would expect that the negative impact of programming languages on maintainability
would be more intense due to the buzz of the last years surrounding the idea that
poor design of programming languages and lack of using the best practices lead to
more buggy and vulnerable code~\cite{Ray:2017:LSP:3144574.3126905, 2019arXiv190110220B}.
However, Figure~\ref{fig:lang_main} shows that only \emph{C} has a negative impact of more than 50\% on software maintainability. We suspect that these values are the result of projects
contributions policies. 9 out of 10 projects --- the equivalent to 31.8\% of the commits ---
of the top-10 of dataset projects with more contributors have restrict contribution
policies regarding code standards. Some examples are the \texttt{laravel/framework},
\texttt{php/php-src}, \texttt{openssl/openssl} and
\texttt{cakephp/cakephp}. For instance, \texttt{laravel/framework} requires the
\texttt{PSR-2 Standard}\footnote{PSR-2 Standard available at
https://www.php-fig.org/psr/psr-2/ (Accessed on \today{})}.

>>>>

In this paper, our concern is to study whether, while improving software
security, developers are also negatively impacting the maintainability
of their software application. To answer the following two
research questions, we use a dataset of security refactorings~\cite{Reis:2017:IJSSE} to measure the
impact of the changes in the maintainability of open-source software.

@@

Projects with \emph{Broken Authentication} flaws are the most affected ones by security
refactorings: 55.3\% (21) of the refactorings had a negative impact in the maintainability
while 35\% (11) yielded a positive impact, and only 13.8\% (6) did not suffer any
change. The results also show that fixing \emph{Memory Leak}s affects the maintainability
of 51.3\% (41) of the cases and leaves only 13.8\% (11) of the cases unaffected.
In 35\% (28) of the cases, maintainability increased. Refactoring vulnerabilities
that allow \emph{Denial-of-Service} attacks (e.g., CVE-2014-3506) have a negative impact
of 47.5\% (19) in the maintainability of open-source software while 30.0\% (12) of
the cases exhibit maintainability improvement and 22.5\% (9) show that
maintainability does not suffer any change. The \emph{Miscellaneous} refactorings pattern
integrates several types of security flaws. These type of refactorings harm maintainability in 43.1\% (84) of the cases. In 31.8\% (62), there is no affect in maintainability and for the other 25.1\% (49) maintainability increases.

It is also important to notice that overall the maintainability of open-source
software is not affected by \emph{Cross-Site Scripting} vulnerabilities and by \emph{Using
Components with Known Vulnerabilities}. \emph{Cross-Site Scripting} does not suffer any
decrease in the maintainability of 55.3\% (63) of cases and 21.1\% (24) have a
positive effect on it. Only 23.7\% (27) of the cases hinder the software
maintainability. From the 21 cases evaluated for \emph{Using Componentes with Known
Vulnerabilities}, only 4.8\% (1) have a positive effect in the maintainability
of software while 14.3\% (3) harms maintainability software, and 81\%(17) have no effect
in the maintainability.

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.57\textwidth]{figures/category.pdf}
 	\caption{Maintainability difference by pattern after security refactorings}
	\label{fig:pat}
\end{figure}

@@

We would expect that the negative impact of programming languages on maintainability
would be more intense due to the buzz of the last years surrounding the idea that
poor design of programming languages and lack of using the best practices lead to
more buggy and vulnerable code~\cite{Ray:2017:LSP:3144574.3126905, 2019arXiv190110220B}.
However, Figure~\ref{fig:lang_main} shows that only \emph{C} has a negative impact of more than 50\% on software maintainability. We suspect that these values are the result of projects
contributions policies. Nine out of ten projects --- the equivalent to 31.8\% of the commits ---
of the top-10 of dataset projects with more contributors have restrict contribution
policies regarding code standards. Some examples are the \texttt{laravel/framework},
\texttt{php/php-src}, \texttt{openssl/openssl} and
\texttt{cakephp/cakephp}. For instance, \texttt{laravel/framework} requires the
\texttt{PSR-2 Standard}\footnote{PSR-2 Standard available at
https://www.php-fig.org/psr/psr-2/ (Accessed on \today{})}.
